\plaintitle

### 2018-06-20 paper.tex
partial correction

Touch gestures are a form of interaction that is used by almost everyone daily in our society. Whether Smartphone, Tablet or Smartwatch - many mobile devices are nowadays controlled by this input technique to master different tasks. High demands are attached to the fast and intuitive use of these devices. The combination of different electronic devices can help to achieve these requirements in a new measure. For example, smart watches are already used to control mobile phones or various home electronics (e.g. controlling lamps). The development of innovative input devices, which can be integrated into clothing or even applied to the user's skin, offers the potential to communicate with electronic devices in completely new ways and make interaction even more intuitive. Especially we focus on touch input devices that are attached to the user's skin or are integrated into clothing and can be used to control different electronic devices.

Because these technologies create new and unusual forms of interaction, we will also look on third-party perceptions of a user’s interactions with a wearable e-textile interface. Social conventions play an important role in the development of these new forms of interaction. The willingness to make gestures depends largely on how appropriate these actions look and feel in public. \cite{touch-wrist} The usability of these technologies must also be considered. To support the user as efficiently, effectively and satisfactorily as possible, the designers must also consider the interaction with the system in addition to the placement of this body-worn controllers.

\section{Showcase of technology examples}
%
% Maybe first a definition of body-worn controllers and a not, that we want to concentrate on Touch-Input.
% I think it is a good idea to pick a few examples which were mentioned in the papers and discuss them briefly.
% For the first part (just the showcase) it is sufficient to just write down the basic technology (with Source): So what it is? How does is work (very briefly)? What it is used for?
% We can get more into detail in the pros/cons section below

In this paper we will analyze some body worn controllers, input sources made joining small electric devices with everyday worn clothing. This allows to have bigger controlling surfaces ready to be used without needing to take out the main device. Moreover, these controllers allows to express some inputs like many users would do in the social interactions, allowing to express emotions while acting the input gesture. This is enabled because the input surface is linked also to the skin, which is a direct input for the user. An example made in More Than Touch \cite[p.185]{more-touch} is “Uncomfortable interactions were chosen for actions that are very important and not reversible”, which shows how the gesture chosen to perform an action were related to the consequence that this action may lead to.

Augmented Reality related research is still trying to optimise the input, which interacts with the shown interface. This research area could potentially capitalize on the natural and unobtrusive ways you interact with Wearables.
“Belt” \cite{belt} is an Body-worn controller specifically build to interact with Augmented Reality applications. As the name suggests the device is worn around the hip, like a regular belt. The device is touch-sensitive and can differentiate between simple gestures and locate the position of the touch-input \cite[p. 2136]{belt}.%([8] Page 2136).
 One of the main use cases of the device are spatial mapped shortcuts to open apps, like a virtual wallet for example. \cite{belt}

\section{Multi-Touch Skin: A thin and flexible Multi touch Sensor for On-Skin Input}
% also true for iSkin
A current approach deals with the placement of multi-touch sensors directly on the skin. Because the human body surface is differently curved and deformable, the system must be particularly flexible and adaptable to fit the skin. This is a special requirement for the design of the sensors as it needs to satisfy the comfort of the user wearing it, who should not feel the fact that he is wearing a sensor, and the necessity for the input to be fixed on a position of the body to recognize correctly the input, which requires tightness onto the skin. \cite{ulbrich2,iSkin}%[2, 3]

Furthermore, the human body has capacitive effects, which must be correctly shielded from an input sensor to ensure reliable touch detection which places special demands on the technology's material: it must both be insulated by the conductive sweat produced by the skin, but it has to let the same sweat flow outside, to avoid the user to feel uncomfortable due to the impossibility of the the skin to dissipate the heat in that area. \cite{ulbrich2} %[2]
To master these challenges, the researchers of “Multi-Touch Skin” \cite{ulbrich2} studied various materials and manufacturing processes. As a result of this research work, a multi-touch sensor was developed, which can be applied directly to the user's skin in form of a thin layer and thus adapts flexibly to the different body shapes. To counteract the capacitive properties of the human body, the prototype consists of three layers, the transmitting, receiving and shielding layer. The individual layers are produced using a standard printer. Silver ink was used here because silver is a particularly effective electrical conductor and so is perfectly qualified to produce the sensor units. Furthermore, this method is well suited to produce multi-touch sensors in different sizes and therefore to test them on different body regions. \cite{ulbrich2}
A small multi-touch sensor behind the ear could be used to manipulate an audio device. For example, the user could adjust the volume by stroking the sensor or pause playback with a simple click. A wristband sized sensor could be used to control interactive lamps. Users could place two fingers around the wrist and rotate them to change the color of a light bulb or adjust the brightness of a lamp using swipe gestures. \cite{ulbrich2}

Another solution is presented by iSkin \cite{iSkin}, where a different material is used to allow input to be sensed. The polydimethylsiloxane (PDMS), the silicon-based material used, is “fully transparent, elastic, and a highly biocompatible material. [...] Therefore it is widely used on or in the human body, for example in body implants.” \cite[p. 2994]{iSkin} PDMS is non-conductive, but adding carbon to it lead to cPDMS, which can create a conductive surface. This allows to create gray sensible surfaces with specific drawn theme on it, which can immediately let the user know what to press for an action, but it doesn’t allow gestures to be performed on, as the principle on which it’s build is an RC charge-discharge time evaluation of the sensible surface.

>>>>

Touch gestures are a form of interaction that is used by almost everyone daily in our society. Whether Smartphone, Tablet or Smartwatch - many mobile devices are nowadays controlled by this input technique to master different tasks. High demands are attached to the fast and intuitive use of these devices. The combination of different electronic devices can help achieving these requirements in a new measure. For example, smart watches are already used to control mobile phones or various home electronics (e.g. controlling lamps). The development of innovative input devices, which can be integrated into clothing or even applied to the user's skin, offers the potential to communicate with electronic devices in completely new ways and make interaction even more intuitive. In this paper we especially focus on touch input devices that are attached to the user's skin or are integrated into clothing and can be used to control different electronic devices.
Because these technologies create new and unusual forms of interaction, we will also look on third-party perceptions of a user’s interactions with a wearable e-textile interface. Social conventions play an important role in the development of these new forms of interaction. The willingness to perform gestures depends largely on how appropriate these actions look and feel in public. \cite{touch-wrist} The usability of these technologies must also be examined: to support the user as efficiently, effectively and satisfactorily as possible, the designers must consider the interaction with the system in addition to the placement of this body-worn controllers.

\section{Showcase of technology examples}
%
% Maybe first a definition of body-worn controllers and a not, that we want to concentrate on Touch-Input.
% I think it is a good idea to pick a few examples which were mentioned in the papers and discuss them briefly.
% For the first part (just the showcase) it is sufficient to just write down the basic technology (with Source): So what it is? How does is work (very briefly)? What it is used for?
% We can get more into detail in the pros/cons section below

The body worn controllers \footnote{input sources made by joining small electric devices with everyday worn clothing} %footnote
allows to have bigger controlling surfaces ready to be used without needing to take out the main device. Moreover, these controllers allow to express input gestures, which are comparable to everyday gestures. Therefore allowing the user to express emotions while using the device. If on a touchscreen the feedback of the input is based only on vibration and visual input, with body-worn controllers the input surface is set over the skin, which allows a direct feedback to the user, that includes the emotion put into performing a gesture. An example made by Wiegel et Al. %in More Than Touch
\cite[p.185]{more-touch} is “Uncomfortable interactions were chosen for actions that are very important and not reversible”, which shows how the gesture chosen to perform an action were related to the consequence that this action may lead to.

%Belt feels a bit off in this position. We should group it in one of the other sections (probably eTextiles) or add a header just for itself, as an example for a body worn controller.
%agree - Andrea

Augmented Reality related research is still trying to optimise the input, which interacts with the shown interface, so that it could potentially capitalize on the natural and unobtrusive ways you interact with Wearables.
“Belt” \cite{belt} is an Body-worn controller specifically build to interact with Augmented Reality applications. As the name suggests the device is worn around the hip, like a regular belt which is touch-sensitive and can differentiate between simple gestures and locate the position of the touch-input \cite[p. 2136]{belt}.%([8] Page 2136).
 One of the main use cases of the device are spatial mapped shortcuts to open apps, like a virtual wallet. \cite{belt}

\section{Multi-Touch Skin: A thin and flexible Multi touch Sensor for On-Skin Input}
% also true for iSkin
A current approach deals with the placement of multi-touch sensors directly on the skin. Because the human body surface is differently curved and deformable, the system must be particularly flexible and adaptable to fit the skin. This is a special requirement for the design of the sensors as it needs to satisfy the comfort of the user wearing it%, who should not feel the fact that he is wearing a sensor,
and the necessity for the input to be fixed on a position of the body to recognize correctly the input, which requires tightness to the skin. \cite{ulbrich2,iSkin}%[2, 3]

Furthermore, the human body has capacitive effects, which must be correctly shielded from an input sensor to ensure reliable touch detection. This places special demands on the technology's material: it must both be insulated by the conductive sweat produced by the skin, but it has to let the same sweat flow outside, to avoid the user to feel uncomfortable due to the impossibility of the the skin to dissipate the heat in that area. \cite{ulbrich2} %[2]
To master these challenges, the researchers of “Multi-Touch Skin” \cite{ulbrich2} studied various materials and manufacturing processes. As a result of this research work, a multi-touch sensor was developed, which can be applied directly to the user's skin in form of a thin layer and thus adapts flexibly to the different body shapes. To counteract the capacitive properties of the human body, the prototype consists of three layers, the transmitting, receiving and shielding layer. The individual layers are produced using a standard printer. Silver ink was used here because silver is a particularly effective electrical conductor and so is perfectly qualified to produce the sensor units. Furthermore, this method is well suited to produce multi-touch sensors in different sizes and therefore to test them on different body regions. \cite{ulbrich2}
A small multi-touch sensor behind the ear could be used to manipulate an audio device. For example, the user could adjust the volume by stroking the sensor or pause playback with a simple click. A wristband sized sensor could be used to control interactive lamps. Users could place two fingers around the wrist and rotate them to change the color of a light bulb or adjust the brightness of a lamp using swipe gestures. \cite{ulbrich2}

Another solution is presented by iSkin \cite{iSkin}, where a different material is used to allow input to be sensed. The polydimethylsiloxane (PDMS), the silicon-based material used, is “fully transparent, elastic, and a highly biocompatible material. [...] Therefore it is widely used on or in the human body, for example in body implants.” \cite[p. 2994]{iSkin} PDMS is non-conductive, but adding carbon to it lead to cPDMS, which can create a conductive surface. This allows to create gray sensible surfaces with specific drawn theme on it, that can immediately let the user know what to press for an action, but it doesn’t allow gestures to be performed on as the principle on which it’s build is an RC charge-discharge time evaluation of the sensible surface \footnote{An RC electronic circuit has a typical charge-discharge time constant $\tau$ which is given by the product $R \cdot C$. when pressure is applied, capacitance changes as the distance between the two surfaces it’s reduced. When a different charging time is sensed, the pressure is identified.}.
